Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2017-03-08 sections0.07/intro0.07.tex
change intro part on how using K -means and what is the different between K-means and SOM

%ANNs
Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

@@

%SOM
A Kohonen self-organizing map (also called self-organizing map, or SOM) is an \textbf{(semi)-}supervised neural network \textbf{\textcolor{red}{for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.}}
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
\textbf{An SOM can be trained unsupervised, and its results can be used to classify other data.
Therefore, this method can be considered as (semi)-supervised.}

@@

%K-means vs. SOM
\textbf{K-means clustering is another unsupervised method that is being used in many astronomical studies \citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-mean clustering divides data into K clusters, in the way that each data belong to the cluster with the nearest mean value.
In the first order, both SOM and K-means work similarly.
In fact, SOM is a combination of K-means and considering the effect of the neighbors which can be ``controlled" by the neighborhood functions (i.e., smoothing process). 
In SOM approach a neighborhood function has an important role in the operation of the SOM. 
When the radius of the neighbourhood function goes to zero, the algorithm will lose its ordering power and will act as k-means clustering.
Because of the smoothing process, the SOM can outperform the K-means clustering even in a simple 1D SOM, which can provide a big picture of clustering similar to the result of K-means clustering.}

@@

%What we did
In this paper, we compare supervised and unsupervised methods of galaxy spectral classification directly, using exactly the same data as \citetalias{Hossein12}.  
Our hypothesis was that since a self-organizing map has the freedom to classify objects in between known classes, it could be applied to this data 
to find the best \textbf{\textcolor{red}{spectral class}} for the 37 galaxies that could not be classified by the supervised method.
First, we train self-organizing maps using the \citetalias{Kinney96} spectral templates and compare our classification with the \citetalias{Kinney96} classification.
Then, we use the trained network to classify \textbf{\textcolor{red}{the spectra}} of 142 galaxies with $0.5 < z < 1$ from \citetalias{Hossein12}. 
As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
\textbf{We cluster both \citetalias{Kinney96} templates and \citetalias{Hossein12} data, usign K-means clustering method and compare the results with the results from SOM clustering.}
In Section~\ref{sec: data_highZ}, we present the data that we used to train our networks and the data classified using the trained networks. 
We describe the SOM and \textbf{K-means clustering methods} in Section~\ref{sec: method_somz}. 
The results of the \textbf{\textcolor{red}{spectral classifications}} and a comparison with previous studies are presented in Section~\ref{sec: result}. 
In Section~\ref{sec: summary_SOMZ}, we summarize our results and discuss potential future work in this subject.

>>>>

%ANNs
Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16a,Hossein16b,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

@@

%SOM
A Kohonen self-organizing map (also called self-organizing map, or SOM) is an \textbf{(semi)-}supervised neural network \textbf{\textcolor{red}{for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.}}
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
\textbf{An SOM can be trained without supervision and its results used to classify other data.
Therefore, this method can be considered as (semi)-supervised.}

@@

%K-means vs. SOM
\textbf{K-means clustering is another unsupervised method used in many astronomical studies \citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-means clustering divides data into $K$ clusters, in a way that each data point belongs to the cluster with the nearest mean value.
To first order, SOM and K-means work similarly.
In fact, SOM is a combination of K-means and considering the effect of the neighbors which can be ``controlled" by the neighborhood functions (i.e., smoothing process). 
In the SOM approach, a neighborhood function has an important role in the operation of the SOM. 
When the radius of the neighbourhood function goes to zero, the algorithm will lose its ordering power and will act like K-means clustering.
Because of the smoothing process, the SOM can outperform K-means clustering even in a simple 1D SOM, which can provide a big picture of clustering similar to the result of K-means.}

@@

%What we did
In this paper, we compare supervised and unsupervised methods of galaxy spectral classification directly, using exactly the same data as \citetalias{Hossein12}.  
Our hypothesis was that since a self-organizing map has the freedom to classify objects in between known classes, it could be applied to this data 
to find the best \textbf{\textcolor{red}{spectral class}} for the 37 galaxies that could not be classified by the supervised method.
First, we train self-organizing maps using the \citetalias{Kinney96} spectral templates and compare our classification with the \citetalias{Kinney96} classification.
Then, we use the trained network to classify \textbf{\textcolor{red}{the spectra}} of 142 galaxies with $0.5 < z < 1$ from \citetalias{Hossein12}. 
As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
\textbf{We cluster both \citetalias{Kinney96} templates and \citetalias{Hossein12} data using the K-means clustering method and compare the results with the results from SOM clustering.}
In Section~\ref{sec: data_highZ}, we present the data that we used to train our networks and the data classified using the trained networks. 
We describe the SOM and \textbf{K-means clustering methods} in Section~\ref{sec: method_somz}. 
The results of the \textbf{\textcolor{red}{spectral classifications}} and a comparison with previous studies are presented in Section~\ref{sec: result}. 
In Section~\ref{sec: summary_SOMZ}, we summarize our results and discuss potential future work in this subject.

### 2017-03-08 sections0.08/data0.08.tex
change intro part on how using K -means and what is the different between K-means and SOM

\section{Data}
\label{sec: data_highZ}
We use \textbf{\textcolor{red}{spectral templates}} from \citetalias{Kinney96}  \textbf{\textcolor{red}{(in 10~\AA\ wavelength bins)}} to train neural networks. 
We classify the \textbf{\textcolor{red}{spectra}} of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12} using the trained networks, and use their physical properties to test the new classifications.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in \textbf{\textcolor{red}{spectral}} clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney Spectral Model}
     \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{images/k96.jpg}
        \caption[\citet{Kinney96} spectral templates for 12 types of galaxies]{\citet{Kinney96} spectral templates for 12 types of galaxies, from fig. 1 in \citet{Hossein12}. The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and Sc show spectra that belong to quiescent galaxies. Starburst galaxy spectra are indicated with SB1 to SB6. Higher numbers represent more intrinsic extinction. \textbf{\textcolor{red}{Note that the vertical axis label from \citet{Hossein12} is incorrect:  the quantity plotted in the vertical axis is flux density, not flux.}}}
        \label{fig: k96}
    \end{figure}

    \citetalias{Kinney96} used ultraviolet-optical spectra of 70 star-forming and quiescent nearby galaxies to produce a set of templates that contained 12 types of \textbf{\textcolor{red}{spectral templates}}.
    These templates have been widely used in many studies to determine morphological type of galaxies or properties of specific types of galaxies~\citep[e.g.][]{Shakouri16, Paiano16, Laporte16, Holden16}.
    \citetalias{Kinney96} stated that these templates can also be used to classify the \textbf{\textcolor{red}{spectra}} of high-redshift galaxies.

    The 12 templates are divided based on their morphological types for quiescent galaxies or their extinction for starburst galaxies (Fig.~\ref{fig: k96}). 
    The quiescent group of galaxies includes Bulge (B), Elliptical (E), S0, Sa, Sb, and Sc galaxies.
    The bulge group represents galaxies similar to M31 and M81, whose UV and optical spectra are dominated by their bulge stellar populations.
    The starburst galaxies are divided into six groups (SB1 to SB6) based on their intrinsic extinctions ($E(B-V)$). 
    As Fig.~\ref{fig: k96} shows, SB1 galaxies have lower internal extinctions ($E(B-V) \simeq 0.05$), while SB6 galaxies have the highest amount of extinction ($E(B-V) \simeq 0.65$) among starburst galaxies. 
    In the quiescent (B to Sb) templates, the spectrum is redder; strong absorption lines and the 4000~\AA~break are distinguishable.
    The SEDs of starburst galaxies are flatter in the optical and near-infrared region than those of the quiescent ones and show strong emission lines.
    For more details on each spectral type, we encourage readers to see \citetalias{Kinney96} and references therein. 
   The \citetalias{Kinney96} spectra span from $\sim1200$~\AA~to $10000$~\AA~with a resolution of $\sim 10$~\AA.
    However, in this work we only use \textbf{\textcolor{red}{observations in the rest-frame wavelength range}} $\sim1200< \lambda < 8000$~\AA~to train our networks; 
    this wavelength range was chosen due to the availability of flux information in those wavelengths for all 12 templates.

 \subsection{SED and Properties of the Sample Galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} was available.
   \citetalias{Hossein12} \textbf{\textcolor{red}{matched the point spread function (PSF) of the photometric data with \citetalias{Kinney96} data and used the photometry as inputs to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
    \citetalias{Hossein12} produced the best SED match, with the wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code.}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and old stellar population with age of $\sim 10$~Gyr, he derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Table~\ref{tab: props}.
    In Section~\ref{sec: 1D_somz}, we study these properties for each category.}}
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

@@

    We classify SEDs that were produced by \citetalias{Hossein12} using the created networks.
    These SEDs are publicly available~\footnote{All the SEDs produced by \citetalias{Hossein12}  can be found at this \href{http://telbib.eso.org/detail.php?bibcode=2012AJ....144..172T}{ESO webpage}.} in the form of flux per rest frame wavelength over a wide range of wavelengths.
    Since we have chosen the \citetalias{Hossein12} SEDs to be classified by the trained network, we only used the part of the SEDs that have the same wavelength range as the \citetalias{Kinney96} \textbf{\textcolor{red}{spectral}} templates.

>>>>

\section{Data}
\label{sec: data_highZ}
We use spectral templates from \citetalias{Kinney96} (in 10~\AA\ wavelength bins) to train neural networks. 
We classify the spectra of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12} using the trained networks, and use their physical properties to test the new classifications.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in spectral clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney Spectral Templates}
     \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{images/k96.jpg}
        \caption[\citet{Kinney96} spectral templates for 12 types of galaxies]{\citet{Kinney96} spectral templates for 12 types of galaxies, from fig. 1 in \citet{Hossein12}. The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and Sc show spectra that belong to quiescent galaxies. Starburst galaxy spectra are indicated with SB1 to SB6. Higher numbers represent more intrinsic extinction. Note that the vertical axis label from \citet{Hossein12} is incorrect:  the quantity plotted in the vertical axis is flux density, not flux.}
        \label{fig: k96}
    \end{figure}

    \citetalias{Kinney96} used ultraviolet-optical spectra of 70 star-forming and quiescent nearby galaxies to produce a set of templates that contained 12 types of spectral templates.
    These templates have been widely used in many studies to determine morphological type of galaxies or properties of specific types of galaxies~\citep[e.g.][]{Shakouri16, Paiano16, Laporte16, Holden16}.
    \citetalias{Kinney96} stated that these templates can also be used to classify the spectra of high-redshift galaxies.

    The 12 templates are divided based on their morphological types for quiescent galaxies or their extinction for starburst galaxies (Fig.~\ref{fig: k96}). 
    The quiescent group of galaxies includes Bulge (B), Elliptical (E), S0, Sa, Sb, and Sc galaxies.
    The bulge group represents galaxies similar to M31 and M81, whose UV and optical spectra are dominated by their bulge stellar populations.
    The starburst galaxies are divided into six groups (SB1 to SB6) based on their intrinsic extinctions ($E(B-V)$). 
    As Fig.~\ref{fig: k96} shows, SB1 galaxies have lower internal extinctions ($E(B-V) \simeq 0.05$), while SB6 galaxies have the highest amount of extinction ($E(B-V) \simeq 0.65$) among starburst galaxies. 
    In the quiescent (B to Sb) templates, the spectrum is redder; strong absorption lines and the 4000~\AA~break are distinguishable.
    The SEDs of starburst galaxies are flatter in the optical and near-infrared region than those of the quiescent ones and show strong emission lines.
    For more details on each spectral type, we encourage readers to see \citetalias{Kinney96} and references therein. 
   The \citetalias{Kinney96} spectra span from $\sim1200$~\AA~to $10000$~\AA~with a resolution of $\sim 10$~\AA.
    However, in this work we only use observations in the rest-frame wavelength range $\sim1200< \lambda < 8000$~\AA~to train our networks; 
    this wavelength range was chosen due to the availability of flux information in those wavelengths for all 12 templates.

 \subsection{SED and Properties of the Sample Galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} were available.
%  We have gone back and forth on the following sentence several times. The commented-out version (what you had) doesn't make sense to me. I replaced it with what I understand was done, based on reading T12.
%   \citetalias{Hossein12} matched the point spread function (PSF) of the photometric data with \citetalias{Kinney96} data 
\citetalias{Hossein12} used point-spread-function matched photometry 
    from \citet{Santini09} as input to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
    \citetalias{Hossein12} produced the best SED match, with the wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code.}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and an old stellar population with age of $\sim 10$~Gyr, he derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Table~\ref{tab: props}.
    In Section~\ref{sec: 1D_somz}, we study these properties for each category.
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

@@

    We classify SEDs that were produced by \citetalias{Hossein12} using the created networks.
    These SEDs are publicly available~\footnote{All the SEDs produced by \citetalias{Hossein12}  can be found at this \href{http://telbib.eso.org/detail.php?bibcode=2012AJ....144..172T}{ESO webpage}.} in the form of flux per rest frame wavelength over a wide range of wavelengths.
    Since we have chosen the \citetalias{Hossein12} SEDs to be classified by the trained network, we only used the part of the SEDs that have the same wavelength range as the \citetalias{Kinney96} spectral templates.

### 2017-03-08 sections0.08/method0.08.tex
change intro part on how using K -means and what is the different between K-means and SOM

\subsection{K-means clustering}
\label{sec: kmeans_method}     
\textbf{In the K-means method, the user decides in advance on a number of clusters, $K$, which cannot be greater than the dimension of data. 
The algorithm randomly chooses $K$ random points to be set as the centroids of the clusters.
Then it finds the data points that are closest to the each assumed centroid and clusters them together.
Mean values of the data points assigned to each cluster are calculated and become the new centroid of the cluster. 
The algorithm finds new clusters based on the new centroids and repeats the procedure until it converges.
We used \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform the K-means clustering.}

>>>>

     %%%After_second_referee_report
     \textbf{As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required.
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     Based on the location of the new vectors on the SOM, they can be compared with the original data set. %20170122PB: unclear what "they" refers to
     Sections~\ref{sec: 1Dv} and ~\ref{sec: 2D} show the results of galaxy spectral classifications using networks trained on template spectra.}

\subsection{K-means clustering}
\label{sec: kmeans_method}     
In the K-means method, the user decides in advance on a number of clusters, $K$, which cannot be greater than the number of data points. 
The algorithm randomly chooses $K$ random points to be the centroids of the clusters.
Then it finds the data points that are closest to each assumed centroid and clusters them together.
The mean values of the data points assigned to each cluster are calculated and become the new centroid of the cluster. 
The algorithm finds new clusters based on the new centroids and repeats the procedure until it converges.
We used \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform the K-means clustering.
